Describe Participants: Characteristics, Diet, and Cardiometabolic Risk

### 2021-11-26 Rmarkdown/analysis/Characteristics_Diet_CardiometabolicRisk_211111.Rmd
Update workflow for paper

```{r Quantile TMAO}
tmao_quantile <- quantile(phen$tmao)
phen$tmao_quantile <- ifelse(phen$tmao <= tmao_quantile[2], "Quantile1",
                             ifelse(phen$tmao > tmao_quantile[2] & phen$tmao <= tmao_quantile[3], "Quantile2",
                                   ifelse(phen$tmao > tmao_quantile[3] & phen$tmao <= tmao_quantile[4], "Quantile3", "Quantile4")))

@@

@@

@@

@@

@@

```{r Meat intake}
# Beef, veal, pork, lamb, and game meat; excludes organ meat and cured meat (oz. eq.)
summary(full$pf_meat_totals)

# Frankfurters, sausages, corned beef, and luncheon meat that are made from beef, pork, or poultry (oz. eq.)
summary(full$pf_curedmeat_totals)

# Total meat, poultry, organ meat, cured meat, seafood, eggs, soy, and nuts and seeds; excludes legumes (oz. eq.)
summary(full$pf_total_totals)

# Total of meat, poultry, seafood, organ meat, and cured meat (oz. eq.)
summary(full$pf_mps_total_totals)

# Kcal
summary(full$kcal_tnfs)
```

>>>>

```{r Quantile TMAO}
# Quantile
tmao_quantile <- quantile(phen$tmao)
phen$tmao_quantile <- ifelse(phen$tmao <= tmao_quantile[2], "Quantile1",
                             ifelse(phen$tmao > tmao_quantile[2] & phen$tmao <= tmao_quantile[3], "Quantile2",
                                   ifelse(phen$tmao > tmao_quantile[3] & phen$tmao <= tmao_quantile[4], "Quantile3", "Quantile4")))

@@

# Tertile
tmao_tertile <- quantile(phen$tmao, c(0:3/3))
phen$tmao_tertile <- ifelse(phen$tmao <= tmao_tertile[2], "Tertile1",
                             ifelse(phen$tmao > tmao_tertile[2] & phen$tmao <= tmao_tertile[3], "Tertile2", "Tertile3"))
# Factor it
phen$tmao_tertile <- factor(phen$tmao_tertile)
table(phen$tmao_tertile)

# Print tertiles for manuscript
tmao_tertile

@@

Save

@@

Make similar tables but for TMAO tertile

```{r Participant characterisitics by TMAO tertile}
# TMAO table by tmao_tertile - females
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.ft

# TMAO table by tmao_tertile - males
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Male",], # note 6103 and 7098 have high il6
                                    factorVars = factorVars)
table_tmao.mt

# TMAO table by tmao_tertile - all, add sex
factorVars <- c("tmao_tertile")
vars <- c("age", "sex", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al", "hdl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.allt

# Do BH correction
allt_ContTable <- attr(table_tmao.allt$ContTable, "pValues")
n <- length(allt_ContTable$pNormal)
# BH correction
allt_ContTable$padj <- stats::p.adjust(allt_ContTable$pNormal, method = "BH", n = n) 
allt_ContTable
```

Save

```{r Save characteristic tertile tables}
table_save.ft <- print(table_tmao.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byTMAOTertile.csv")

table_save.mt <- print(table_tmao.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byTMAOTertile.csv")

table_save.allt <- print(table_tmao.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile.csv")

# BH corrections for the all tertile table
write.csv(allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile_Padj.csv")
```

@@

TMAO tertile 
Make TMAO food intake by tertile table
```{r Food Intake by TMAO tertile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "hei_asa24_totalveg",
          "hei_asa24_sfat",
          "hei_asa24_",
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.ft

# TMAO table by age bin - males
table_food.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Male",],
                                    factorVars = factorVars)
table_food.mt

# TMAO table by sex
table_food.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full,
                                    factorVars = factorVars)
table_food.allt

# Do BH correction
food_allt_ContTable <- attr(table_food.allt$ContTable, "pValues")
n <- length(food_allt_ContTable$pNormal)
# BH correction
food_allt_ContTable$padj <- stats::p.adjust(food_allt_ContTable$pNormal, method = "BH", n = n) 
food_allt_ContTable
```

```{r Save food tertile tables}
table_save.ft <- print(table_food.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Food_Female_byTMAOTertile.csv")

table_save.mt <- print(table_food.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Food_Male_byTMAOTertile.csv")

table_save.allt <- print(table_food.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile.csv")

write.csv(food_allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile_PadjTable.csv")
```

@@

```{r Meat intake}
# Beef, veal, pork, lamb, and game meat; excludes organ meat and cured meat (oz. eq.)
summary(full$pf_meat_totals)
sd(full$pf_meat_totals)

# Frankfurters, sausages, corned beef, and luncheon meat that are made from beef, pork, or poultry (oz. eq.)
summary(full$pf_curedmeat_totals)
sd(full$pf_curedmeat_totals)

# Chicken, turkey, Cornish hens, duck, goose, quail, and pheasant (game birds); excludes organ meat and cured meat (oz. eq.)
summary(full$pf_poult_totals)
sd(full$pf_poult_totals)

# Total meat, poultry, organ meat, cured meat, seafood, eggs, soy, and nuts and seeds; excludes legumes (oz. eq.)
summary(full$pf_total_totals)
sd(full$pf_total_totals)

# Total of meat, poultry, seafood, organ meat, and cured meat (oz. eq.)
summary(full$pf_mps_total_totals)
sd(full$pf_mps_total_totals)

# Eggs (chicken, duck, goose, quail) and egg substitutes (oz. eq.)
summary(full$pf_eggs_totals)
sd(full$pf_eggs_totals)

# HEI total score
summary(full$hei_asa24_totalscore)
sd(full$hei_asa24_totalscore, na.rm = TRUE)

# Kcal
summary(full$kcal_tnfs)
sd(full$kcal_tnfs)
```
